---
title: "Lab 1 - Data visualization"
author: "Siqi Lan"
format: pdf
editor: visual
editor_options: 
  chunk_output_type: console
---

## Load Packages

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(viridis)
```
```{r}
#| label: load-data
data(midwest)
view(midwest)
```

## Exercise 1

(Type your answer to Exercise 1 here. Add code chunks as needed. Don't forget to label your code chunk. Do not use spaces in code chunk labels.)

```{r}
#| label: histogram-ex1
#| message: false
ggplot(data = midwest,
       aes(x = popdensity)) +
  geom_histogram(binwidth = 10000) +
  labs(x = "Population density (person/unit area)", 
       y = "Number of Counties", 
       title = "The Population Density of Counties")
```
1) Describe the shape of the distribution.
  The distribution is right skewed.
2) Does there appear to be any outliers? Briefly explain.
  Yes, from the histgram we can detect that there are some extremely high density around 87500 people per area.

## Exercise 2

(Type your answer to Exercise 2 here. Add code chunks as needed. Don't forget to label your code chunk. Do not use spaces in code chunk labels.)

```{r}
#| label: histogram-ex2 
#| message: false
ggplot(data = midwest,
       aes(x = percollege, y = percbelowpoverty, color = state)) +
  geom_point() +
  labs(x = "Percentage of People with a College Degree", 
       y = "Percentage below Poverty", 
       title = "Percentage of People with a College Degree \nVersus 
Percentage below Poverty") +
  scale_color_viridis_d(option = "A", end = 0.8) +
  theme_minimal()
```

## Exercise 3

(Type your answer to Exercise 3 here. Add code chunks as needed. Don't forget to label your code chunk. Do not use spaces in code chunk labels.)

```{r}

```

## Exercise 4

```{r}

```

## Exercise 5

```{r}

```

## Exercise 6

```{r}

```

## Exercise 7

```{r}

```
